State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-13 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
loads of little edits and typos, up to line 285. Dinner time!

\section{Introduction} This paper presents a theoretical framework for generating macroscopic neural field mathematical models from patient-specific data. There has been much interest in generating physiologically plausible neural field models to study brain dynamics at the meso/macroscopic scale. While our understanding of the function of neurons is well developed, the overall behaviour of the brain's meso and macro-scale dynamics remains largely a theoretical. Understanding the brain at this level is extremely important since this is the scale where pathologies such as epilepsy, Parkinson's disease and schizophrenia are manifested. Mathematical neural field models provide insights into the underlying physics and dynamics of electroencephalography (EEG) and magnetoencephalography (MEG) (see \cite{Deco2008,David2003} for recent reviews). These models have demonstrated possible mechanisms for the genesis of neural rhythms (such as the alpha and gamma rhythms) \cite{Liley1999,RENNIE2000}, epileptic seizure generation \cite{DaSilva2003,Suffczynski2004,Wendling2005} and insights into other pathologies \cite{Moran2008,Schiff2009} that would be difficult to gain from experimental data alone.

@@

For models to be clinically viable they must be patient specific. A possible approach to achieve this would be to fit a general continuum neural field model, like the Wilson and Cohen (WC) or Amari~\cite{Wilson1973,Amari1977} style model or a neural mass model like the Jansen and Ritt model \cite{Jansen1995}, to patient specific EEG data. Fitting the neural models to individuals is a highly non-trivial task, and until very recently this has not been reported in the literature.

@@

This paper extends the work of Schiff and Sauer~\cite{schiff2008kalman} by establishing a framework for estimating the state of the WC equations for higher order systems via a systematic model reduction procedure. In addition, a method is presented for estimating the connectivity structure and the synaptic time constant. Until now, estimation of local intracortical connectivity structure has not been reported (to the best of the authors knowledge). Recently, it has been shown that it is possible to estimate local coupling of spatiotemporal systems using techniques from control systems theory and machine learning \cite{Dewar2009}. The key development in that work was to represent the spatiotemporal system as a standard state-space model with an order independent of the number of observations (recording electrodes in this case). In addition, the appropriate model selection tools have been developed \cite{Scerri2009} allowing for the application of the technique to neural fields.

Modelling the neural dynamics within this framework has the distinct advantage over the more standard multivariate auto-regressive (AR) models: that the number of parameters to define the spatial connectivity is considerably smaller than the number of AR coefficients typically required to achieve the required model complexity.

@@

\section{Neural Field Model}\label{NeuralModelSection} In this section, we describe a variant of the WC or Amari style neural field model \cite{Wilson1973,Amari1977} that was used to create our intracortical connectivity estimator. Neural field models relate mean firing rates of pre-synaptic neural populations to mean post-synaptic membrane potentials. They are popular due to being parsimonious yet having a strong link with the underlying physiology. Each neural population represents a functional cortical processing unit, such as a column. The columnar organisation of the cortex is continuous, where pyramidal cells are members of many columns. In general, cortical structure can be modelled in a physiologically plausible manner as being locally homogeneous (in short range intracortical connectivity) and heterogeneous (in long range cortico-cortical and corticothalamic connectivity)~\cite{Jirsa2009,Qubbaj2007}. In certain regoins of the cortex, each column is thought be connected locally via symmetric short range local excitation, with surround inhibition \cite{Braitenberg1998}. For example, this structural organisation is most studied in the visual system, where the surrounding inhibition effectively tunes a cortical column to a particular receptive visual field~\cite{Sullivan2006}. Neural field models are descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour~\cite{David2003,Bressloff2001,Breakspear2006}. 
Field models are also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations~\cite{Amari1977,Coombes2005,Coombes2007}.

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameters $f_{max}$ and $v_0$ describe the maximum firing rate and firing threshold of the neural populations. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) &=&  \int_{-\infty}^t h\left(t - t'\right) \\
	&\times&\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}'dt'. \nonumber
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} &+& \zeta v\left( \mathbf{r},t \right) = \\
	&&\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. \nonumber
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretise time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) &=& \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \nonumber\\ 
	&+& e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is i.i.d. such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$. This term was added to account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma_{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be considered a decomposition into a set of basis functions where 

@@

	\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 

@@

		c_{n_y,L} & \dots & c_{n_y,L} 

@@

where $L$ is the number of basis functions and 

@@

There are two options available to alleviate problems with aliasing. The first, which is the most desirable, is to increase the spatial sampling frequency (using a higher resolution of sensors). The second is to use wider sensors, where the spatial averaging acts like a low-pass anti-aliasing filter. This option is a compromise between alleviating issues with aliasing and losing high frequency information contained in the field. In practice, it is difficult  to estimate the bandwidth of the cortex using traditional electrophysiological measurements due to spatial limitations. However, we envisage it may be possible the estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency components contribute negligibly to the inverse problem and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results. \omg{Although our results are severely effected by the loss of high frequency info! Ken: Do you think you can word this a little better than what I have done.}

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}. 
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian
\begin{equation}\label{eq:nDimensionalGaussian}
	\phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
 is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving \omg{need to rewrite eq~\ref{eq:WidthFrequencyRelationship} in terms of $\sigma_{\phi}$. At the mo it is writen in terms of $\sigma_{nu}$}
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.
The dynamic field must be spatially oversampled using a larger number of sensors then imposed by frequency analysis considerations, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors, although this is not a requirement for using this estimation framework. \omg{Ken: can you please write something about what should be done in this case?} For a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Following this, the knowledge of the distance between basis functions directly implies the total number of basis functions required.

% 
% Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section we describe the procedure for estimating the states, $\mathbf{x}_t$, that scale the field basis functions and the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic time constant, $\xi$. The estimation process is a two part iterative algorithm, consisting of the state estimation step followed by the parameter estimation step. At each iteration, the sequence of estimated states is used to update the parameter set, then the resulting parameters are used to estimate a new state vector for the next iteration. The procedure stops when the parameters converge. The algorithm is initialised using a bounded random state sequence which guarantees the initial estimated parameter set forms a stable kernel. \omg{Parham: can you please add to this to include the initialisation of the field?}

The unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS uses the unscented Kalman Filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows

>>>>

\section{Introduction} This paper presents a theoretical framework for generating macroscopic neural field mathematical models from patient-specific data. There has been much interest in generating physiologically plausible neural field models to study brain dynamics at the meso/macroscopic scale. While our understanding of the function of neurons is well developed, the overall behaviour of the brain's meso and macro-scale dynamics remains largely theoretical. Understanding the brain at this level is extremely important since it is this scale where pathologies such as epilepsy, Parkinson's disease and schizophrenia are manifested. Mathematical neural field models provide insights into the underlying physics and dynamics of electroencephalography (EEG) and magnetoencephalography (MEG) (see \cite{Deco2008,David2003} for recent reviews). These models have demonstrated possible mechanisms for the genesis of neural rhythms (such as the alpha and gamma rhythms) \cite{Liley1999,RENNIE2000}, epileptic seizure generation \cite{DaSilva2003,Suffczynski2004,Wendling2005} and insights into other pathologies \cite{Moran2008,Schiff2009} that would be difficult to gain from experimental data alone.

@@

For models to be clinically viable they must be patient specific. A possible approach to achieve this would be to fit a general continuum neural field model, like the Wilson and Cohen (WC) or Amari~\cite{Wilson1973,Amari1977}  models or a neural mass model like the Jansen and Ritt model \cite{Jansen1995}, to patient specific EEG data. Fitting the neural models to individuals is a highly non-trivial task, and until very recently this has not been reported in the literature.

@@

This paper extends the work of Schiff and Sauer~\cite{schiff2008kalman} by establishing a framework for estimating the state of the WC equations for higher order systems via a systematic model reduction procedure. In addition, a method is presented for estimating the connectivity structure and the synaptic time constant. Until now, estimation of local intracortical connectivity structure has not been reported (to the best of the authors knowledge). Recently, it has been shown that it is possible to estimate local coupling of spatiotemporal systems using techniques from control systems theory and machine learning \cite{Dewar2009}. The key development here was to represent the spatiotemporal system as a standard state-space model with an order independent of the number of observation locations (recording electrodes in this case). In addition, the appropriate model selection tools have been developed \cite{Scerri2009} allowing for the application of the technique to neural fields.

Modelling the neural dynamics within this framework has a distinct advantage over the more standard multivariate auto-regressive (AR) models: that the number of parameters to define the spatial connectivity is considerably smaller than the number of AR coefficients typically required to achieve the required model complexity.

@@

\section{Neural Field Model}\label{NeuralModelSection} In this section, we describe a variant of the WC neural field model that was used to create our intracortical connectivity estimator. Neural field models relate mean firing rates of pre-synaptic neural populations to mean post-synaptic membrane potentials. They are popular due to being parsimonious yet having a strong link with the underlying physiology. Each neural population represents a functional cortical processing unit, such as a column. The columnar organisation of the cortex is continuous, where pyramidal cells are members of many columns. In general, cortical structure can be modelled in a physiologically plausible manner as being locally homogeneous (in short range intracortical connectivity) and heterogeneous (in long range cortico-cortical and corticothalamic connectivity)~\cite{Jirsa2009,Qubbaj2007}. In certain regions of the cortex, each column is thought to be connected locally via symmetric short range local excitation, with surround inhibition \cite{Braitenberg1998}. For example, this structural organisation is most studied in the visual system, where the surrounding inhibition effectively tunes a cortical column to a particular receptive visual field~\cite{Sullivan2006}. Neural field models are descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour~\cite{David2003,Bressloff2001,Breakspear2006}. Field models are also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations~\cite{Amari1977,Coombes2005,Coombes2007}.

@@

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameters $f_{max}$ and $v_0$ describe the maximum firing rate and firing threshold of the neural populations. The parameter $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into \ref{SpikesToPotential} we get the spatiotemporal model 
\begin{eqnarray}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) &=&  \int_{-\infty}^t h\left(t - t'\right) \\
	&\times&\int_\Omega w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}'dt'. \nonumber
\end{eqnarray}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} Dh\left( t \right) = \delta \left( t \right), 
\end{equation}
where $D=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{eqnarray}
	\label{FinalFormContinuous} \frac{dv\left( \mathbf{r},t \right)}{dt} &+& \zeta v\left( \mathbf{r},t \right) = \\
	&&\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. \nonumber
\end{eqnarray}
To arrive at the integro-difference equation (IDE) form of the model we discretise time using a first-order Euler method (see~\ref{Time Discretization}) giving 
\begin{eqnarray}
	\label{DiscreteTimeModel} v_{t+1}\left(\mathbf{r}\right) &=& \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} \nonumber\\ 
	&+& e_t\left(\mathbf{r}\right), 
\end{eqnarray}
where $T_s$ is the time step, $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is i.i.d. such that $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$ denotes a spatial Gaussian Process with mean zero and covariance function $\gamma(\mathbf{s}-\mathbf{r})$ \cite{Rasmussen2006} This term was added to account for uncertainty in the model. The mapping between the membrane voltage and the collected iEEG/LFP data is modelled using the observation function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field, $n=1,...,N$ indexes the sensors, $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\Sigma_{\varepsilon}\right)$ and $\mathcal{N}\left(0,\Sigma_{\varepsilon}\right)$ denotes the multivariate Normal distribution with mean zero and covariance matrix $\Sigma_{\varepsilon}$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry where 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}. 
\end{equation}

@@

where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The choice of Gaussian basis functions can be justified by the existence of the so called bump solutions for this class of model, which have a Gaussian shape~\cite{Coombes2005}. The width and positioning of the basis functions can be determined by spectral analysis (explained in detail in Section~\ref{SpectralAnalysisSection}). The connectivity kernel can also be decomposed into a set of basis functions where 

@@

	\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{-1} 

@@

		c_{n_y,1} & \dots & c_{n_y,L} 

@@

where $L$ is the number of basis functions used to decompose the voltage field and 

@@

There are two options available to alleviate problems with aliasing. The first, which is the most desirable, is to increase the spatial sampling frequency (using a higher resolution of sensors). The second is to use wider sensors, where the spatial averaging acts like a low-pass anti-aliasing filter. This option is a compromise between alleviating issues with aliasing and losing high frequency information contained in the field. In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements due to spatial limitations. However, we envisage it may be possible the estimate the bandwidth of the cortex using other modalities with higher spatial resolution such as fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Errors associated with mild aliasing result in distortions in the high frequency components of the field dynamics. However, after applying the basis function decomposition, high frequency components contribute negligibly to the inverse problem and are therefore not a major problem for the state and parameter estimation~\cite{Sanner1992}. This is further illustrated our results. \omg{Although our results are severely effected by the loss of high frequency info! Ken: Do you think you can word this a little better than what I have done.}

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions must also satisfy Shannon's sampling theorem to capture the dominant spectral information of the field. Thus, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_b < \frac{1}{2\rho_b\nu_{co}}
\end{equation}
where $\rho_b$ is an over-sampling parameter to determine the basis function separation and $\nu_{co}$ is the spatial cut-off frequency from the observations. For the Gaussian basis functions, $\phi(\mathbf{r})$, being considered, the basis function width parameter, $\sigma_{\phi}^2$, can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian
\begin{equation}\label{eq:nDimensionalGaussian}
	\phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
 is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{co}$ giving \omg{need to rewrite eq~\ref{eq:WidthFrequencyRelationship} in terms of $\sigma_{\phi}$. At the mo it is writen in terms of $\sigma_{nu}$}
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{co}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{co}}= \frac{2\boldsymbol\nu_{co}^\top \boldsymbol\nu_{co}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent a field with frequency content up to $\boldsymbol\nu_{co}$. Proofs for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.
The dynamic field must be spatially oversampled using a larger number of sensors than imposed by the frequency analysis, to use frequency analysis to determine the minimum distance and width of the basis functions. Following this, the distance between basis functions is usually larger then the distance between sensors, although this is not a requirement for using this estimation framework. \omg{Ken: can you please write something about what should be done in this case?} For a spatially isotropic field, the basis functions can be placed on a regular grid and hence the distance between basis functions dictates the total number of basis functions required.

% 
% Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section we describe the procedure for estimating the states, $\mathbf{x}_t$, and the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic time constant, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. At each iteration, the sequence of estimated states is used to update the parameter set, then the resulting parameters are used to estimate a new state vector for the next iteration. The procedure stops when the parameters converge. The algorithm is initialised using a bounded random state sequence which guarantees the initial estimated parameter set forms a stable kernel. \omg{Parham: can you please add to this to include the initialisation of the field?}

The unscented Rauch Tung Striebel Smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman Filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows
